Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-11-17 paper/sec-pileup.tex
deleted old plots
work on PU section

\section{Robustness under realistic pile-up conditions}

@@

The discussion so far has not included the effects of
pile-up, under realistic conditions of the HL-LHC.
%
In this section we show how our qualitative results, in particular
the enhanced discrimination provided by the MVA, are robust
even when realistic pile-up conditions are accounted for.
%
First of all we describe the simulation of realistic
pile-up conditions at the HL-LHC, and the corresponding
subtraction using {\tt SoftKiller}.
%
Then we present the validation of the pile-up subtraction,
and compare a number of distributions, including substructure variables,
with and without PU.
%
Finally we revisit the MVA analysis of Sect.~\ref{sec:mva}, and
show that our qualitative conclusions are unchanged even
in the presence of realistic PU conditions.

\subsection{Pile-up subtraction with {\tt SoftKiller}}

With this motivation, a large number
of minimum bias events have been added to the various signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios for the amount of pile-up expected
at the HL-LHC, one with a mean number of
PU vertices of $\la n_{\rm PU}\ra=80$, and another
with $\la n_{\rm PU}\ra=150$.
%
Minimum bias events have been generated with {\tt Pythia8} using
the Monash 2013 tunes.
%
We have generated a total of 100M minimum-bias events, which are
then superimposed to the hard-scattering events in a random way.

In order to subtract the pile-up, a number of techniques
have been developed recently. While PU will be be one of the major problems at the HL-LHC, fortunately
a number of powerful methods for the subtraction of  UE, MPI and PU
effects 
have been developed recently~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba,Ellis:2009me,Bertolini:2014bba,Cacciari:2014gra,Cacciari:2014jta,Berta:2014eza,Larkoski:2014wba}.
%
While a complete optimization of the PU subtraction strategy is beyond
the scope of this work, here we will use the {\tt SoftKiller}
method~\cite{Cacciari:2014gra}
as implemented in {\tt FastJet}.

The idea underlying {\tt SoftKiller} is based on eliminating particles
below a given cut-off in transverse momentum, $p_T^{\rm cut}$, whose
value is dynamically determined in a way that makes the event-wide
transverse-momentum flow density $\rho$
\be
\rho={\rm median} \Bigg\{ \frac{p_{Ti}}{A_i}\Bigg\}
\ee
where the median is computed over all the patches $i$ with area
$A_i$ and transverse momentum $p_{Ti}$ in which the event is partitioned.
%
The size and number of these patches is a free parameter of the algorithm -
in this work we will use a value $a=0.4$ for the length of these
patches.
%
We restrict ourselves to the rapidity interval
$\eta \in \lc -2.5, 2.5\rc$ for the estimation of the
$p_T$ flow density $\rho$, since only the central region
is used in this analysis.
%
In this work, the {\tt SoftKiller} method is applied
to the particles at the end of the parton shower, before
they are clustered into the anti-$k_t$ jets.

\subsection{Validation of pile-up subtraction}

Now we turn to validate our PU subtraction strategy.
%
To this end, we compare various kinematical distributions in three
cases:
\begin{itemize}
\item without any pile-up,
\item with pile-up, but without any PU subtraction,
  \item with pile-up and with {\tt SoftKiller} subtraction.
\end{itemize}
As we will show now, some substructure variables in the boosted
regime are particularly robust against PU effects, and this
will have the important consequence that the MVA discrimination
will not be substantially affected.

In Fig.~\ref{fig: pu-vs-nopu-comparison-d2} we
compare the distribution of the $D_2^{(\beta)}$
    substructure variable, Eq.~(\ref{eq:d2}), for the
    boosted category in two cases: without PU, and
    with $\la n_{\rm PU}\ra=150$ PU events.
    %
    In the latter case, PU has been subtracted using
    {\tt SoftKiller}.
    %
    We see that the distribution is relatively unaffected by PU.
    %
    Note that by construction $D_2^{(\beta)}$ has been designed
    to be insensitive to PU, so this agreement is an important
    cross-check.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
  \includegraphics[width=0.60\textwidth]{plots/pu-vs-nopu-comparison-d2.pdf}
  \caption{\small
    Comparison of the distribution of the $D_2^{(\beta)}$
    substructure variable, Eq.~(\ref{eq:d2}), for the
    boosted category in two cases: without PU, and
    with $\la n_{\rm PU}\ra=150$ PU events.
    %
    In the latter case, PU has been subtracted using
    {\tt SoftKiller}.
    %
    See text for more details.
}
\label{fig:pu-vs-nopu-comparison-d2}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

We now compare a number of signal distributions with
and without PU.
%
In Fig.~\ref{fig:m_H_PU} we show first of all the invariant mass distribution
of the leading and subleading Higgs candidates in the boosted
and resolved categories.
%
These distributions are plotted after the $b$-tagging,
before they are used as input to the MVA.
%
As we can see, in the boosted category, the residual effects of PU
after the SoftKiller subtraction are very mild in this
distribution.
%
The effects are more important for the resolved category, where PU
shifts the Higgs peak by am amount $\Delta m_h \simeq 5$ GeV.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
      \includegraphics[width=0.49\textwidth]{plots/m_H0_res_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H1_res_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H0_bst_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H1_bst_comp.pdf}
  \caption{\small
    Comparison of the invariant mass distributions of the leading (left plots)
    and subleading (right plots) Higgs candidates in the resolved
    (upper plots) and boosted (lower plots) categories,
    without PU and with $\la n_{PU}\ra=80$ subtracted with SoftKiller.
}
\label{fig:m_H_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\section{Impact of pile-up}

@@

The discussion so far has not accounted for the effects of
pile-up (PU), which are known to be very important at the HL-LHC.
%
In this section, we show how our qualitative results, in particular
the enhanced discrimination provided by the MVA, are robust
in the presence of realistic PU conditions.

First of all we describe the simulation of realistic
PU conditions at the HL-LHC, and discuss the settings of
the PU subtraction strategy that we adopt.
%
Then we present the validation of the PU subtraction,
and compare a number of distributions, including substructure variables,
with and without PU.
%
Finally we revisit the MVA analysis of Sect.~\ref{sec:mva}, and
show that our qualitative conclusions are robust
in the presence of realistic PU conditions.
%
As we will show now, a signal significance for
Higgs production in the $4b$ final state
of up to $S/\sqrt{B}\simeq 4$
can be achieved at the HL-LHC even accounting for PU.

\subsection{PU subtraction with {\tt SoftKiller}}

To study the impact of PU in our analysis,
a large number
of Minimum Bias (MB) events,
including Multiple Parton Interactions (MPI),
have been added to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios for the amount of PU expected
at the HL-LHC, one with a mean number of
PU vertices of $\la n_{\rm PU}\ra=80$, and another
with $\la n_{\rm PU}\ra=150$.
%
MB events have been generated with {\tt Pythia8} using
the Monash 2013 tune.
%
We have generated a total of 100M minimum-bias events, which are
then randomly superimposed to the hard-scattering events.

In order to subtract the PU, a large number of techniques
have been proposed recently~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba,Ellis:2009me,Bertolini:2014bba,Cacciari:2014gra,Cacciari:2014jta,Berta:2014eza,Larkoski:2014wba}.
%
In this work, PU  will be subtracted by means
of the the {\tt SoftKiller}
method~\cite{Cacciari:2014gra}, as implemented in {\tt FastJet}.
%
Let us note however that
a systematic optimization of the PU subtraction strategy is beyond
the scope of this work, and in particular that further
improvements over the results presented below is feasible
with further tuning of the PU subtraction.

The idea underlying {\tt SoftKiller} is based on eliminating particles
below a given cut-off in their transverse momentum, $p_T^{\rm (cut)}$, whose
value is dynamically determined in a way that makes the event-wide
transverse-momentum flow density $\rho$ vanish.
%
This $p_T$ flow density is defined as
\be
\rho\equiv{\rm median}_i \Bigg\{ \frac{p_{Ti}}{A_i}\Bigg\} \, ,
\ee
where the median is computed over all the patches $i$ with area
$A_i$ and transverse momentum $p_{Ti}$ in which the $\lp \eta,\phi\rp$ plane
is partitioned.
%
From its definition, we observe that the value of $p_T^{(\rm cut)}$
will be dynamically raised until half of the patches have $\rho=0$.

The size and number of these patches is a free parameter of the algorithm -
here we will use square patched with length $a=0.4$.
%
We restrict ourselves to the central rapidity interval,
$\eta \in \lc -2.5, 2.5\rc$, for the estimation of the
$p_T$ flow density $\rho$, which is the region relevant
for the present analysis.
%
In this work, the {\tt SoftKiller} method is applied
to particles at the end of the parton shower, before
the jet clustering stage.

\subsection{Validation of the PU subtraction}

Now we turn to validate our PU subtraction strategy.
%
To this end, we compare different kinematical distributions
in the three relevant cases:
cases:
\begin{itemize}
\item without any PU,
\item adding PU, but without any PU subtraction,
  \item adding PU, followed by the {\tt SoftKiller} subtraction.
\end{itemize}
As we will show now, {\tt SoftKiller} exhibits a reasonable
performance in PU subtraction.
%
We will also see that the boosted category is less affected
by PU than the resolved category, as expected from the higher
$p_T$ threshold in the corresponding selection cuts.

Let us begin by comparing
a number of signal distributions with
and without PU, with $\la n_{\rm PU}\ra=80$ in the former
case.
%
In Fig.~\ref{fig:m_H_PU} we show the invariant mass distribution
of the leading and subleading Higgs candidates. in the boosted
and resolved categories.
%
These distributions are plotted after the $b$-tagging, that is,
before they are used as input to the MVA.
%
As we can see, in the boosted category, the residual effects of PU
after the {\tt SoftKiller} subtraction are very mild in these
distributions, with the position of the Higgs mass peaks essentially
unchanged.
%
The effects are more important for the resolved category, where PU
shifts the Higgs peak by an amount $\Delta m_h \simeq 5$ GeV.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
      \includegraphics[width=0.49\textwidth]{plots/m_H0_res_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H1_res_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H0_bst_comp.pdf}
      \includegraphics[width=0.49\textwidth]{plots/m_H1_bst_comp.pdf}
  \caption{\small
    Comparison of the invariant mass distributions of the leading (left plots)
    and subleading (right plots) Higgs candidates in the resolved
    (upper plots) and boosted (lower plots) categories,
    without PU and with $\la n_{PU}\ra=80$ subtracted with {\tt SoftKiller}.
}
\label{fig:m_H_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Next we compare the transverse momentum of the leading Higgs
candidate, $p_t^{h_1}$ and the invariant mass of the di-Higgs system
$m_{hh}$, in Fig.~\ref{fig:mHH_PU}, both for the boosted and
for the resolved categories.
%
In the case of the $p_T$, the differences between the resolved
and boosted selection cuts can be clearly observed.
%
The effect of PU is negligible in the boosted case, and small
in the resolved case, except for large $p_T$ values.
%
For the case of the $m_{hh}$ distribution, similar conclusions
apply.
%
Note that this latter observable is particularly sensitive
of BSM dynamics.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \vspace{-1cm}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/pt_H0_C2_bst_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_res_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/m_HH_C2_bst_comp.pdf}
  \caption{\small
    Comparison of the transverse momentum $p_T^h$ of the leading
    Higgs candidate (upper plots) and of the invariant mass $m_{hh}$
    of the di-Higgs system (lower plots) in the resolved
    (left plots) and boosted (right plots) categories,
    without PU and with $\la n_{PU}\ra=80$ subtracted with {\tt SoftKiller}.
}
\label{fig:mHH_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Then we assess the impact of PU in some of the substructure variables
used as input to the MVA in the boosted category.
%
In particular we consider the subjetiness variable,
$\tau_{21}$, and the ratio
of energy correlation functions $D_2^{(\beta)}$,
corresponding to the the leading Higgs candidate.
%
This comparison is illustrated in Fig.~\ref{fig:Substructure_PU}.
%
As can be seen, these substructure variables, which
as demonstrated before carry a very substantial
discrimination power, are relatively unaffected by PU.
%
Therefore, we don't expect a substantial loss of discrimination
power due to PU effects in our analysis.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \vspace{-1cm}
  \includegraphics[width=0.49\textwidth]{plots/D2_h0_bst_comp.pdf}
  \includegraphics[width=0.49\textwidth]{plots/tau21_h0_bst_comp.pdf}
   \caption{\small
     Comparison of the substructure variables $D_2^{(\beta)}$ (left)
     and $\tau_{21}$ (right)
     for the leading Higgs candidate in the boosted category,
   without PU and with $\la n_{PU}\ra=80$ subtracted with {\tt SoftKiller}.
}
\label{fig:Substructure_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

Up to know we have restricted ourselves to the study of the impact of PU
on signal distributions.
%
We now assess how the signal over background discrimination can be affected
in the presence of realistic PU conditions.

{\bf add plots signal vs background with PU}
